---
title: "P600_across_L1_and_L2"
author: "Ioannis Iliopoulos"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
## Introduction

Here I analyze the data I collected from a within-subject ERP study on a grammatical violation.  
The same individuals were tested in the same grammatical violation in their L1 (German) and their L2 (English).  
Our aim is to identify:  
(a) at the group-level whether the expected P600 is elicited in both L1 and L2,  
(b) whether third factors modulate the P600, and  
(c) whether individual variability is consistent across L1 and L2.  

Here I do (a) and (b). For (c), I tried extracting the random slopes by participant and then doing a correlation analysis — but maybe there is a more efficient way to do this.

## Load Required Libraries

```{r}
library(lme4)
library(lmerTest)  # extends lme4 with p-values
library(Matrix)
library(ggplot2)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(dplyr)
library(tidyr)
library(RColorBrewer)
```

## 1. Read Data
The data include the following varibles:
Subject : Subject/Participant Number
Condition: Grammatical - Ungrammatical
Item_Nr: Item Number
Language: German (DE) - English (EN)
value: Amplitude (our dependent variable)
OQPT_Score: L2 prodificiency Score 
WM_Score : Working Memory Score 
VF_Score : Verbal Fluency Score 

```{r}
merged_df <- read.csv("merged_P600_data.csv")
```

## 2. Prepare Variables

```{r}
merged_df <- merged_df %>%
  mutate(
    language = factor(language),
    condition = factor(condition),
    WM_Score = scale(as.numeric(WM_Score)),
    proficiency_centered = scale(as.numeric(OQPT_Score)),
    VF_Score = scale(as.numeric(VF_Score))
  )

contrasts(merged_df$language) <- contr.sum(2) / -2
contrasts(merged_df$condition) <- contr.sum(2) / -2
```

## 3. Main Mixed Models

```{r}
m1i <- lmer(
  value ~ language * condition + 
    (1 + condition + language | subject) + 
    (1 + condition | item_n),
  data = merged_df
)

m1iWM <- lmer(
  value ~ WM_Score * language * condition + 
    (1 | subject) + 
    (1 | item_n),
  data = merged_df
)

round(summary(m1i)$coef, 3)
round(summary(m1iWM)$coef, 3)
```

## 4. Subset L1 (German)

```{r}
merged_df_DE <- merged_df %>%
  filter(language == "DE") %>%
  droplevels()

m1iWM_DE <- lmer(
  value ~ WM_Score * condition + 
    (1 + condition | subject) + 
    (1 | item_n),
  data = merged_df_DE
)

round(summary(m1iWM_DE)$coef, 3)
```

## 5. Subset L2 (English)

```{r}
merged_df_EN <- merged_df %>%
  filter(language == "EN") %>%
  droplevels()

levels(merged_df_EN$condition) <- c("grammatical", "ungrammatical")
contrasts(merged_df_EN$condition) <- contr.sum(2) / -2
```

## 6. L2 Predictors

```{r}
# Proficiency model
m1i_prof <- lmer(
  value ~ proficiency_centered * condition +
    (1 | subject) +
    (1 + condition | item_n),
  data = merged_df_EN
)

# Verbal Fluency model
m1i_VF <- lmer(
  value ~ VF_Score * condition +
    (1 | subject) +
    (1 | item_n),
  data = merged_df_EN
)

# WM model
m1i_WM <- lmer(
  value ~ WM_Score * condition +
    (1 | subject) +
    (1 + condition | item_n),
  data = merged_df_EN
)

round(summary(m1i_prof)$coef, 3)
round(summary(m1i_VF)$coef, 3)
round(summary(m1i_WM)$coef, 3)
```
