---
title: "RePsychLing Ioannis Iliopoulos's 'P600_across_L1_and_L2'"
author: "Reinhold Kliegl"
date: today
date-format: iso
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-fold: false
    number-sections: true
    fig-width: 8
    fig-height: 6
    fig-format: svg
editor_options: 
  chunk_output_type: console
---

# Introduction

Here I analyze the data I collected from a within-subject ERP study on a grammatical violation.  
The same individuals were tested in the same grammatical violation in their L1 (German) and their L2 (English).  
Our aim is to identify:  
(a) at the group-level whether the expected P600 is elicited in both L1 and L2,  
(b) whether third factors modulate the P600, and  
(c) whether individual variability is consistent across L1 and L2.  

Here I do (a) and (b). For (c), I tried extracting the random slopes by participant and then doing a correlation analysis — but maybe there is a more efficient way to do this.

RK: I am not a friend of correlating conditional modes extracted from LMM fits and using inferential statistics, because conditional modes by definition are not independent. Perhaps estimate effects within languages and then a correlation parameters for the effects?

# Load required libraries

```{r}
library(arrow)
library(easystats)
library(tidyverse)
library(lme4)
library(summarytools)
```

# Read Data
The data include the following varibles:
Subject : Subject/Participant Number
Condition:  Ungrammatical - Grammatical
Item_Nr: Item Number
Language: English (EN) - German (DE)
value: Amplitude (our dependent variable)
OQPT_Score: L2 prodificiency Score 
WM_Score : Working Memory Score 
VF_Score : Verbal Fluency Score 

# Preprocessing

RK: 

+ You should take care of missing values during preprocessing, that is use the same data across different LMMs; I leave `WM_score` out of the analyses and drop rows with missing in other relevant variables.
+ Do you want to use z-score covariates (or only mean-centered ones)?
+ There is a bit of problematic imnplication for estimates of their interaction
with `contr.sum(2) / -2` and `contr.sum(2) / -2`. 

```{r}
dat <- 
  read.csv("./data/merged_P600_data.csv") |> 
  mutate(
    Subj = as_factor(paste0("S", str_pad(subject, width = 3, side = "left", pad = "0"))),
    Item = as_factor(paste0("I", str_pad(item_n, width = 3, side = "left", pad = "0"))),
    Language = factor(language),
    Condition = factor(condition),
    l2 = scale(as.numeric(OQPT_Score)),
    vf = scale(as.numeric(VF_Score))
  ) |> 
  select(Subj, Item, Language, Condition, l2, vf, value) |> 
  drop_na()

#stview(dfSummary(dat))

contrasts(dat$Language) <- contr.sum(2) 
contrasts(dat$Condition) <- (-1)*contr.sum(2) 
```

```{r}
dat <- 
  read.csv("./data/merged_P600_data.csv") |> 
  mutate(
    Subj = as_factor(paste0("S", str_pad(subject, width = 3, side = "left", pad = "0"))),
    Item = as_factor(paste0("I", str_pad(item_n, width = 3, side = "left", pad = "0"))),
    Language = factor(language),
    Condition = factor(condition),
    l2 = scale(as.numeric(OQPT_Score)),
    vf = scale(as.numeric(VF_Score))
  ) |> 
  select(Subj, Item, Language, Condition, l2, vf, value) |> 
  drop_na()

#stview(dfSummary(dat))

contrasts(dat$Language) <- contr.sum(2) 
contrasts(dat$Condition) <- (-1)*contr.sum(2) 
```

Or use a subset with WM score.

```{r}
#| eval: false
dat_wm <- 
  read.csv("./data/merged_P600_data.csv") |> 
  mutate(
    Subj = as_factor(paste0("S", str_pad(subject, width = 3, side = "left", pad = "0"))),
    Item = as_factor(paste0("I", str_pad(item_n, width = 3, side = "left", pad = "0"))),
    Language = factor(language),
    Condition = factor(condition),
    l2 = scale(as.numeric(OQPT_Score)),
    vf = scale(as.numeric(VF_Score)),
    wm = scale(as.numeric(WM_Score))
  ) |> 
  select(Subj, Item, Language, Condition, l2, vf, wm, value) |> 
  drop_na()

stview(dfSummary(dat_wm))

contrasts(dat_wm$Language) <- contr.sum(2) 
contrasts(dat_wm$Condition) <- (-1)*contr.sum(2) 
```


# Main LMMs

```{r}
system.time(
m1i <- lmer(
  value ~ 1 + Language * Condition + 
         (1 + Condition * Language | Subj) +  (1 + Condition | Item),
  data = dat, REML=FALSE, 
  control=lmerControl(calc.derivs=FALSE))
)

system.time(
m1i_vf_l2 <-  lmer(
  value ~ 1 + (vf+l2) * Language * Condition + 
         (1 + Condition * Language | Subj) +  (1 + Condition | Item),
  data = dat, REML=FALSE, 
  control=lmerControl(calc.derivs=FALSE))
)

anova(m1i, m1i_vf_l2)

round(summary(m1i)$coef, 3)
round(summary(m1i_vf_l2)$coef, 3)
```

No evidence for the relevance of `vf` and `proficiency`.

# Estimating condition effects within DE and within EN

```{r}
system.time(
m2i <- lmer(
  value ~ 1 + Language / Condition + 
         (0 + Language / Condition  | Subj) +  (1 + Condition | Item),
  data = dat, REML=FALSE, 
  control=lmerControl(calc.derivs=FALSE))
)

VarCorr(m2i)
```

Lots of positive correlations. 

+ Means of P600 between DE and EN correlate 0.537
+ Condition effects of DE and EN correlate 0.291

# Save the current dataframe

This file is used as input in Julia MixedModels.jl script **ii_1.qmd**

```{r}
write_feather(data.frame(dat), "./data/P600.arrow")
```

# Main LMMs with WM-score covariate

```{r}
system.time(
m3i <- lmer(
  value ~ 1 + Language * Condition + 
         (1 + Condition * Language | Subj) +  (1  | Item),
  data = dat_wm, REML=FALSE, 
  control=lmerControl(calc.derivs=FALSE))
)

system.time(
m3i_wm <-  lmer(
  value ~ 1 + wm * Language * Condition + 
         (1 + Condition * Language | Subj) +  (1 | Item),
  data = dat_wm, REML=FALSE, 
  control=lmerControl(calc.derivs=FALSE))
)

anova(m3i, m3i_wm)

round(summary(m3i)$coef, 3)
round(summary(m3i_wm)$coef, 3)
```

+ Removed VC Item-related Condition that generated: " boundary (singular) fit: see help('isSingular')"
+ No evidence for the relevance of `wm`.

# Appendix

```{r}
sessionInfo()
```

